---
title: Introduction to `HerringData`
subtitle: '`r read.dcf(file = "../DESCRIPTION", fields = "Title")`'
author: Matthew Grinnell
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
      rmarkdown::html_vignette:
        fig_caption: yes
        toc: true
        number_sections: true
vignette: >
  %\VignetteIndexEntry{Introduction to `HerringData`}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

---

```{r options, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Background

`r read.dcf(file = "../DESCRIPTION", fields = "Description")`

# Set up

Load required packages.

```{r setup, message = FALSE}
library(HerringData)
library(SpawnIndex)
library(tidyverse)
library(DBI)
library(scales)
```

Load required data.

```{r data}
data(pars)
data(codes_group)
data(codes_period)
data(conv_factors)
data(database_info)
```

Set parameters.

```{r pars}
region <- "PRD"
substock <- "StatArea"
yrs_plot <- c(pars$years$assess, 2024)
```

This introduction shows an example workflow in the
`r region` stock assessment region (SAR) using
`r substock` for sub-stock structure.

# Data

The following process for the three data types is similar.
First, load the data and wrangle it into a user-friendly format
via the respective function in the `loaders` family.
Then, filter to the `r region` SAR.
Finally, aggregate and wrangle the data to use in SISCAH
via the respective function in the `SISCAH` family.

## Biological data

First, load and wrangle biological data: number- and weight-at-age.
Then, filter to only include data from the `r region` SAR.
Finally, aggregate the data by year, gear (i.e., period), and stock
to input into SISCAH.
Note that the `structure` argument (set to `r substock` in this example)
indicates the implied stock structure.

```{r bio}
bio <- load_bio(db_info = database_info, quiet = TRUE) %>%
  filter(Region == region) %>%
  siscah_bio(structure = substock)
```

### Number-at-age data

Number-at-age by stock, period, and year for fish aged
`r formals(siscah_bio)$age_min_number`
(i.e., column 'a`r formals(siscah_bio)$age_min_number`') to
`r formals(siscah_bio)$age_max`,
where the age-10 class is a 'plus group' which includes fish ages 10 and older.
The 'Stock' column is the sub-stock name;
in this case it indicates the Statistical Area (e.g., "StatArea03").
The 'Area' column is a number that increments from 1
to the number of sub-stocks.
Note that Gear and Period are synonymous.
The 'Period' column indicates the fishery type:
'Other' is the reduction, the food and bait, as well as the special use fishery;
'RoeSN' is the roe seine fishery; and
'RoeGN' is the roe gillnet fishery.

```{r num_age}
num_age <- bio$number_age  %>%
  left_join(y = codes_period, by = "Gear") %>%
  select(Stock, Area, Period, Year, starts_with("a", ignore.case = FALSE)) %>%
  mutate(Period = factor(Period, levels = codes_period$Period))
num_age
```

### Weight-at-age data

Weight-at-age in kilograms for fish aged
`r formals(siscah_bio)$age_min_weight` to
`r formals(siscah_bio)$age_max`.
Note that table layout is similar to the number-at-age table.
There are two weight-at-age tables;
the first one is weight-at-age by year, period, and stock

```{r wt_age}
wt_age <- bio$weight_age %>%
  left_join(y = codes_period, by = "Gear") %>%
  select(Stock, Area, Period, Year, starts_with("a", ignore.case = FALSE)) %>%
  mutate(Period = factor(Period, levels = codes_period$Period))
wt_age
```

The second one is weight-at-age by year and stock for seine gear only.

```{r wt_age_sn}
wt_age_sn <- bio$weight_age_seine
wt_age_sn
```

## Catch data

Like biological data above,
load and wrangle catch data,
filter to only include data from the `r region` SAR, and
aggregate the data by year, period, and stock to input into SISCAH.

```{r catch}
catch <- load_catch(db_info = database_info, quiet = TRUE) %>%
  filter(Region == region) %>%
  siscah_catch(structure = substock) %>%
  left_join(y = codes_period, by = "Gear") %>%
  select(Stock, Area, Period, Year, Catch) %>%
  mutate(Period = factor(Period, levels = codes_period$Period))
```

Catch by stock, period, and year in thousands of tonnes.
Note that table layout is similar to the biological data tables.

```{r catch_catch}
catch
```

## Spawn data

Like biological and catch data above,
load and wrangle spawn index data,
filter to only include data from the `r region` SAR, and
aggregate the data by year and stock to input into SISCAH.
Note that the spawn index is calculated when the data is loaded
using functions in the `SpawnIndex` package.
SISCAH accounts for two types of spawn index data:
data from surface survey observations and
data from dive survey observations.

```{r spawn}
spawn <- load_spawn(db_info = database_info, quiet = TRUE) %>%
  filter(Region == region) %>%
  siscah_spawn(structure = substock)
```

Spawn index by stock and year in thousands of tonnes.
Note that table layout is similar to the biological and catch data tables.

```{r spawn_spawn}
spawn
```

# Figures

Example figures to display biological, catch, and spawn index data.

## Biological data

Biological data includes number- and weight-at-age.

### Number-at-age data

Plot annual number of fish sampled for each stock and period.

```{r num_age_plot, fig.width = 8, fig.height = 5}
num_age_cap <- paste(
  region, "number of fish sampled from",
  paste(yrs_plot, collapse = " to "), "by stock and period."
)
num_age_long <- num_age %>%
  pivot_longer(
      cols = starts_with("a", ignore.case = FALSE), names_to = "Age",
      values_to = "Number"
    ) %>%
  mutate(Age = parse_number(Age)) 
num_age_plot <- ggplot(
  data = num_age_long, mapping = aes(x = Year, y = Number)
) +
  geom_col() +
  scale_y_continuous(labels = label_comma()) +
  labs(y = "Number of fish") +
  expand_limits(x = yrs_plot) +
  facet_grid(Stock ~ Period) +
  theme_bw()
num_age_plot
```

Plot annual proportion-at-age for each stock and period.

```{r prop_age_plot, fig.width = 8, fig.height = 6}
prop_age_cap <- paste(
  region, "proportion-at-age from",
  paste(yrs_plot, collapse = " to "), "by stock and period."
)
prop_age_long <- num_age_long %>%
  group_by(Stock, Area, Period, Year) %>%
  mutate(Total = sum(Number)) %>%
  ungroup() %>%
  mutate(Proportion = Number / Total) %>%
  filter(Proportion != 0)
prop_age_plot <- ggplot(
  data = prop_age_long, mapping = aes(x = Year, y = Age, size = Proportion)
) +
  geom_point(alpha = 0.6, shape = 16) +
  scale_y_continuous(breaks = pretty_breaks()) +
  scale_size(range = c(0.5, 3)) +
  expand_limits(x = yrs_plot) +
  facet_grid(Stock ~ Period) +
  theme_bw() +
  theme(legend.position = "top")
prop_age_plot
```

### Weight-at-age data

Plot annual weight-at-age for each stock and period.

```{r wt_age_plot, fig.width = 8, fig.height = 6}
wt_age_cap <- paste(
  region, "weight-at-age in kilograms (kg) from",
  paste(yrs_plot, collapse = " to "), "by stock and period."
)
wt_age_long <- wt_age %>%
  pivot_longer(
      cols = starts_with("a", ignore.case = FALSE), names_to = "Age",
      values_to = "Weight"
    ) %>%
  mutate(Age = parse_number(Age)) 
wt_age_plot <- ggplot(
  data = wt_age_long,
  mapping = aes(x = Year, y = Weight, group = Age, colour = Age)
) +
  geom_line(linewidth = 1) +
  scale_colour_viridis_c() +
  labs(y = "Weight (kg)") +
  expand_limits(x = yrs_plot) +
  facet_grid(Stock ~ Period) +
  theme_bw() +
  theme(legend.position = "top")
wt_age_plot
```

Plot annual weight-at-age for each stock (seine gear only).

```{r wt_age_sn_plot, fig.width = 8, fig.height = 6}
wt_age_sn_cap <- paste(
  region, "weight-at-age in kilograms (kg) from",
  paste(yrs_plot, collapse = " to "), "by stock for seine gear only."
)
wt_age_sn_long <- wt_age_sn %>%
  pivot_longer(
      cols = starts_with("a", ignore.case = FALSE), names_to = "Age",
      values_to = "Weight"
    ) %>%
  mutate(Age = parse_number(Age)) 
wt_age_sn_plot <- ggplot(
  data = wt_age_sn_long,
  mapping = aes(x = Year, y = Weight, group = Age, colour = Age)
) +
  geom_line(linewidth = 1) +
  scale_colour_viridis_c() +
  labs(y = "Weight (kg)") +
  expand_limits(x = yrs_plot) +
  facet_grid(Stock ~ .) +
  theme_bw() +
  theme(legend.position = "top")
wt_age_sn_plot
```

## Catch data

Plot annual catch for each stock and period.

```{r catch_plot, fig.width = 8, fig.height = 6}
catch_cap <- paste(
  region, "catch in thousands of tonnes (t) from",
  paste(yrs_plot, collapse = " to "), "by stock and period."
)
catch_plot <- ggplot(
  data = catch, mapping = aes(x = Year, y = Catch, fill = Period)
) +
  geom_col() +
  scale_fill_viridis_d() +
  labs(y = "Catch (1,000 t)") +
  expand_limits(x = yrs_plot) +
  facet_grid(Stock ~ .) +
  theme_bw() +
  theme(legend.position = "top")
catch_plot
```

## Spawn index data

Pivot the spawn data to facilitate plotting and then 
plot annual spawn index for each stock and survey.

```{r spawn_plot, fig.width = 8, fig.height = 6}
spawn_cap <- paste(
  region, "spawn index in thousands of tonnes (t) from",
  paste(yrs_plot, collapse = " to "), "by stock and survey."
)
spawn_long <- spawn %>%
    pivot_longer(
      cols = c(Surface, Dive), names_to = "Survey", values_to = "Spawn"
    ) %>%
  mutate(Survey = factor(Survey, levels = c("Surface", "Dive")))
spawn_plot <- ggplot(
  data = spawn_long, mapping = aes(x = Year, y = Spawn, fill = Survey)
) +
  geom_col() +
  scale_fill_viridis_d() +
  labs(y = "Spawn index (1,000 t)") +
  expand_limits(x = yrs_plot) +
  facet_grid(Stock ~ .) +
  theme_bw() +
  theme(legend.position = "top")
spawn_plot
```
